---
title: "Piru Enhancement - data preparation"
author: "Shane Dewees"
date: "2022-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here) # similar to set working directory
library(tidyverse) # data wrangling
library(lubridate) # dates and times
library(readr)
```

# MAFA percent cover
```{r mafa_cover, include=TRUE}
cover_import <- read_csv("data/mafa_cover.csv", 
    col_types = cols(
      date = col_character(), 
      plot_number = col_integer(), 
      cover_percent = col_number(), 
      notes = col_character()
      )
    )

mafa_cover <- cover_import %>%
  # filter(plot_type == "enhancement") %>% # removed this column
  select(date, plot_number, MAFA_removal, cover_type, cover_percent) %>% 
  # rename(cover_percent = cover_.) %>% # renamed csv
  rename(treatment = MAFA_removal) %>% 
  mutate(date = mdy(date),
         month = month(date),
         year = year(date)) %>% 
  filter(cover_type %in% c("MAFA", "MAFA alive")) %>%
  mutate_at(vars(cover_percent), funs(round(., 0))) %>%  # rounds cover_percent to integer
  mutate(
    cover_percent = as.integer(cover_percent),
    month = as.integer(month),
    year = as.integer(year)
  ) %>% 
  select(!date) 
  

print(mafa_cover)
```
##Dates of data collection
``` {r mafa cover dates,  include=FALSE}
mafa_cover_dates <- mafa_cover %>% 
  group_by(year, month) %>% 
 dplyr::summarize(mafa_cover_dates = n_distinct(year))

print(mafa_cover_dates) # lists the years and months of data collection
# two monitoring dates: Fall 2017 and spring/summer 2019
# each monitoring event spanned two months 
# 
# 2017	8		
# 2017	9			
# 2019	5			
# 2019	6	
```

# Seedling survival
```{r survivial, include=TRUE} 

################ CAN WE CREATE AN IF ELSE STATEMENT TO CHANGE DATE TO NUMBER OF MONTHS POST TRANSPLANTING?????

survival <- read.csv(here("data", "survival_data.csv")) %>% 
  select(date, plot_number, MAFA_removal, species, alive) %>% 
  rename(treatment = MAFA_removal) %>% 
  mutate(date = mdy(date),
         month = month(date),
         year = year(date),
         treatment = str_remove(treatment, "removal")
        ) %>% 
  group_by(year, month, plot_number, treatment, species) %>% 
  summarise(alive = sum(alive, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(
    month = as.integer(month),
    year = as.integer(year)
  )

print(survival) # will only print the first 100 pages
# view(survival) # full data set for 2018-2021 here


```
```{r seedling survival dates, include = FALSE}
survival_dates <- survival %>% 
  group_by(year, month) %>% 
 dplyr::summarize(survival_dates = n_distinct(year))

print(survival_dates) # lists the years and months of data collection
#  2018 jan planted, feb, march, dec
#  2019 jan planted, march, april, may, june, sept, nov 
#  PAR --> 2019 jan, march, may, july, sept, nov
#  2020 march, may, dec
#  2021 feb, april
```
# Seedling mortality
```{r mortality, include=TRUE}
mortality <- read.csv(here("data", "survival_data.csv")) %>% 
  select(date, plot_number, MAFA_removal, species, dead) %>% 
  rename(treatment = MAFA_removal) %>% 
  mutate(date = mdy(date),
         month = month(date),
         year = year(date),
         treatment = str_remove(treatment, "removal")) %>% 
  group_by(year, month, plot_number, treatment, species) %>% 
  summarise(dead = sum(dead, na.rm = TRUE)) %>% 
    ungroup() %>% 
   mutate(
    month = as.integer(month),
    year = as.integer(year)
  )

print(mortality)
# mortality is the inverse of survival

```
```{r seedling mortality data collection dates, include=FALSE}

mortality_dates <- mortality %>% 
  group_by(year, month) %>% 
 dplyr::summarize(mortality_dates = n_distinct(year))

print(mortality_dates) # lists the years and months of data collection
# mortality_dates should be the same as survival_dates
```


# PAR - sunshine
```{r par, include=TRUE}
par <- read.csv(here("data", "par_data.csv")) %>% 
  select(date, plot, weeding, readingno, ambient_PAR_1, below_PAR_1) %>% 
  rename(plot_number = plot) %>% 
  rename(treatment = weeding) %>% 
  mutate(date = case_when(date == "Jan. 30, 2019"~"1/30/2019",
                          date == "25-Jan-19"~"1/25/2019",
                          date == "24-Mar-19"~"3/24/2019",
                          date == "Nov. 16, 2019" ~ "11/16/2019",
                          date == "9-Sep-19" ~ "9/09/2019",
                          date == "19-Jul-19"~ "7/19/2019",
                          date == "24-May-19" ~ "5/24/2019",
                          TRUE ~ date),
         plot_number = str_remove(plot_number, "ENH0"),
         plot_number = str_remove(plot_number, "ENH"),
         plot_number = as.integer(plot_number),
         treatment = str_remove(treatment, "removal"),
         date = mdy(date)) %>% 
  group_by(date, plot_number, treatment) %>% 
  summarize(ambient_par = mean(ambient_PAR_1),
            below_par = mean(below_PAR_1)) %>% 
  ungroup() %>% 
  mutate(normalized_par = below_par/ambient_par,
         month = month(date),
         year = year(date),
         year = as.integer(year),
         month = as.integer(month)) %>% 
  select(year, month, plot_number, treatment, normalized_par) 

print(par)
```

```{r par data collection dates, include=FALSE}
par_dates <- par %>% 
  group_by(year, month) %>% 
 dplyr::summarize(par_dates = n_distinct(year))

print(par_dates) # lists the years and months of data collection
# 2018 april
# 2019 jan, march, may, july, sept, nov
```


# TDR - soil moisture
```{r tdr, include=TRUE}
tdr <- read.csv(here("data", "tdr_data.csv")) %>% 
  select(date, plot, weeding, reading, TDR.value) %>% 
  rename(plot_number = plot,
         soil_moisture = TDR.value,
         treatment = weeding) %>% 
  mutate(plot_number = str_remove(plot_number, "ENH0"),
         plot_number = str_remove(plot_number, "ENH"),
         plot_number = as.numeric(plot_number),
         treatment = str_remove(treatment, "removal"),
         date = dmy(date)) %>% 
  group_by(date, plot_number, treatment) %>% 
  summarize(soil_moisture = mean(soil_moisture, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(month = month(date),
         year = year(date)) %>% 
  select(!date) %>% 
  select(year, month, plot_number, treatment, soil_moisture) %>% 
  mutate(
    year = as.integer(year),
    month = as.integer(month), 
    plot_number = as.integer(plot_number)
    )
  

print(tdr)
```
```{r tdr dates, include=FALSE}
tdr_dates <- tdr %>% 
  group_by(year, month) %>% 
 dplyr::summarize(tdr_dates = n_distinct(year))

print(tdr_dates) # lists the years and months of data collection
# 2018 april, dec
# 2019 march
# 2020 march, april, may
```


# Join soil moisture (TDR) to Seedling survival 
Join by month, year, and plot number
Purpose: soil moisture as an explainitory variable of seedling survival
```{r survival_soil_moisture, include=TRUE}
survival_soil_moisture <- survival %>% 
  left_join(tdr, by = c("month", "year", "plot_number")) %>% 
  drop_na()

print(survival_soil_moisture)
```


# Join MAFA percent cover and PAR (sunlight)
Join 2019 data by plot number and month 
Purpose: MAFA cover as an explanatory variable of PAR (sunlight) experienced by the seedlings
```{r par_mafacover join, include=TRUE}
mafa_cover.2019 <- mafa_cover %>% 
  filter(year == "2019")

par_mafacover <- par %>% 
  filter(year == "2019") %>% 
  left_join(mafa_cover.2019, by = c("plot_number")) %>% 
  rename(year.PAR = year.x) %>% 
  rename(month.PAR = month.x) %>% 
  rename(year.MAFAcover = year.y) %>% 
  rename(month.MAFAcover = month.y) %>% 
  select(year.PAR, month.PAR, plot_number, normalized_par, cover_percent, treatment.y) %>% 
  drop_na() %>% 
  rename(treatment = treatment.y)

print(par_mafacover)
```


# Join MAFA percent cover and seedling survival in 2018
Join 
Purpose: MAFA cover as an explanatory variable of seedling survival in 2018
```{r par_mafacover join, include=TRUE}
# join 2018 survival data with 2017 cover data
survival_mafacover.2018 <- survival %>% 
  filter(year == "2018") %>% 
  left_join(mafa_cover, by = c("plot_number")) %>% 
  filter (year.y == "2017") %>% 
  rename(year.survival = year.x) %>% 
  rename(month.survival = month.x) %>% 
  rename(year.MAFAcover = year.y) %>% 
  rename(month.MAFAcover = month.y) %>% 
  select(year.survival, month.survival, plot_number, species, alive, cover_percent, year.MAFAcover) %>% 
  drop_na() 

print(survival_mafacover.2018)
```

Purpose: MAFA cover as an explanatory variable of seedling survival in 2019
```{r}
# join 2019 survival data with 2019 cover data

survival_mafacover.2019 <- survival %>% 
  filter(year == "2019") %>% 
  left_join(mafa_cover, by = c("plot_number")) %>% 
  filter (year.y == "2019") %>% 
  rename(year.survival = year.x) %>% 
  rename(month.survival = month.x) %>% 
  rename(year.MAFAcover = year.y) %>% 
  rename(month.MAFAcover = month.y) %>% 
  select(year.survival, month.survival, plot_number, alive, species, cover_percent, year.MAFAcover) %>% 
  drop_na() 

print(survival_mafacover.2019)
```

