---
title: "Piru_enchancement-Models"
author: "Kit Swift"
date: "2022-12-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here) # similar to set working directory
library(tidyverse) # data wrangling
library(lubridate) # dates and times
library(readr)
library(ggthemes)
library(plotly)
library(dplyr)
library(olsrr)
library(survival)
library(lme4)
library(lmerTest)
library(multcomp)
library(multcompView)

devtools::install_github("an-bui/calecopal")

library(calecopal)
# all palettes
names(cal_palettes)

# Site colors: 1, 2 
cal_palette(name = "chaparral1", n = 2, type = "continuous") # [1] "#DCC27A" "#F19B34"

# Treatment colors: Full, half, none 
cal_palette(name = "chaparral3", n = 3, type = "continuous") # [1] "#D3E3CA", "#92A587", "#2F3525"
cal_palette(name = "desert", n = 3, type = "continuous") # [1] "#F6EECF", "#B09175", "#291611"

#mafa_cover_all = MAFA % cover 2018 and 2019
#exotic_cover_all = exotic % cover 2018 and 2019


```


# Question 1: How do environmental conditions change by a) canopy treatment and b) canopy cover?
Is canopy treatment or canopy cover a stronger predictor of the environmental conditions below the shrub canopy? 

Notes:
* __response variable__: The variable you want to test. (e.g.,. PAR, TDR, nncover)
* __explanatory variables__: The variables that can explain the outcome. (e.g., treatment, mafacover, site)
* __data frame__: Include all the data you'll be working with. (e.g., combine par and mafa_cover_all)
* __formula__: a formula is a two-sided linear formula object describing both the _fixed-effects_ and _random effects_ part of the model, with the response on the left of a ~ operator and the terms, separated by + operators, on the right. 
* __random effects__: Random-effects terms are distinguished by vertical bars __(|)__ separating expressions for design matrices from grouping factors. Two vertical bars __(||)__ can be used to specify multiple uncorrelated random effects for the same grouping variable. 
  + Because of the way it is implemented, the ||-syntax works only for design matrices containing numeric (continuous) predictors; to fit models with independent categorical effects, see dummy or the lmer_alt function from the afex package.)

### Load PAR, TDR, and non-native percent cover data.
### Save as a csv in the "Processed" data folder.
```{r load data, include=FALSE}

mafa_cover_all <- read_csv("data/processed/mafa_cover_all.csv")
par <- read_csv("data/processed/par.csv")
tdr <- read_csv("data/processed/tdr.csv")
exotic_cover_all <- read_csv("data/processed/exotic_cover_all.csv")
survmort <- read_csv("data/processed/survmort.csv")
# names(survmort)
# [1] "...1"         "year"         "month"        "plot_number"  "treatment"    "species"     
# [7] "alive"        "dead"         "site"         "month_count"  "transplanted"

# PAR
parmod <- par %>%
  left_join(mafa_cover_all, by = c("year", "site", "plot_number"), na.rm = F) %>% 
  rename(mafacover = cover_percent)
print(parmod)
write.csv(parmod, "data/processed/par_model.csv")
par1 <- filter(parmod, site == "1")
par2 <- filter(parmod, site == "2")

# TDR
tdrmod <- tdr %>% 
  left_join(mafa_cover_all, by = c("year", "site", "plot_number"), na.rm = F) %>% 
  rename(mafacover = cover_percent) %>% 
  filter(year <= 2019)
print(tdrmod)
write.csv(tdrmod, "data/processed/tdr_model.csv")
tdr1 <- filter(tdrmod, site == "1")
tdr2 <- filter(tdrmod, site == "2")

# NN COVER
nnmod <- exotic_cover_all %>% 
   left_join(mafa_cover_all, by = c("year", "site", "plot_number"), na.rm = F) %>% 
  rename(nncover = cover_percent.x,
         mafacover = cover_percent.y, 
         treatment = MAFA_removal) %>% 
  select("date", "year", "site", "plot_number", "treatment", "nncover", "mafacover")
print(nnmod)
write.csv(nnmod, "data/processed/nn_model.csv")
nnmod1 <- filter(nnmod, site == "1")
nnmod2 <- filter(nnmod, site == "2")
```

***

## PAR linear mixed-effects model

1. lmer

2. testing the residuals for normality

```{r par model - normality, include = TRUE}

lme.parmod <- lmer(normalized_par ~ treatment + mafacover + site + month + (1|date) +(1|plot_number) + (1|year),
                   data = parmod)

MOD <- lme.parmod

# Calculate model's residual and fitted values
F1 <- fitted(MOD)
E1 <- residuals(MOD, "pearson")

# See residuals as a histogram, look for normal distribution
hist(E1)

# Graph residuals against a normal distribution, look for all points to be on 1:1 line
#need to call next two lines together
qqnorm(E1, xlab = "Theoretical Quantiles", ylab = "Pearson Residuals")
qqline(E1)

# Graph residuals vs fitted values to inspect homogeneity of variance, look for a random scattering of plots
# run together 2 lines
plot(x=F1, y=E1, xlab="Fitted Model Values", ylab="Pearson Residuals", main = "Variance")
abline(h = 0) # looking for random scatter

# Check for crazy outliers, look for lines MUCH taller than the rest of data
plot(cooks.distance(MOD))

# Plot a box plot for each factor, look for equal variance amount levels along the zero line
# Box plots need to be assessed for every fixed effect (categorical factor) for your model. 
# Random effects need the qqline just like normality. 
# So, only site and treatment get box plots. 
# MAFA cover is a covariate (continuous variable).
boxplot(E1 ~ treatment, data = parmod)
abline(h = 0)

boxplot(E1 ~ month, data = parmod)
abline(h = 0)

boxplot(E1 ~ site, data = parmod)
abline(h = 0)

# Plot each random effect in the model - things in the (1|_), look for dots along 1:1 line
qqnorm(ranef(MOD)$date[,1], main = "Plot Residuals")
qqline(ranef(MOD)$date[,1]) # unique dates

qqnorm(ranef(MOD)$plot_number[,1], main = "Plot Residuals")
qqline(ranef(MOD)$plot_number[,1]) # unique plots
```

3. statistics

```{r par model - stats, include = TRUE}
# print 
print(lme.parmod)
summary(lme.parmod) # p-values
anova(lme.parmod) # fixed level effects, pvalues, etc.
```

4. figures

```{r par model - figures, include=TRUE}

# scatterplot: PAR x mafacover (%)
ggplot(data = parmod, aes(x = mafacover, y = normalized_par)) + 
  geom_point() +
labs(title = "Normalized PAR by MAFA cover (%)", x = "MAFA canopy cover (%)", y = "PAR") +
  geom_smooth(method="lm")+
  theme_bw()

# scatterplot: PAR x mafacover x treatment
ggplot(data = parmod, aes(x = mafacover, y = normalized_par)) + 
  geom_point(aes(color = treatment)) +
  labs(title = "Normalized PAR by MAFA cover (%) and Treatment", x = "MAFA canopy cover (%)", y = "PAR") +
  theme_bw()

# scatterplot: PAR x mafacover x site
ggplot(data = parmod, aes(x = mafacover, y = normalized_par)) + 
  geom_point(aes(color = site)) +
  labs(title = "Normalized PAR by MAFA cover (%) and Site", x = "MAFA canopy cover (%)", y = "PAR") +
  theme_bw() 

```

```{r par model - additional figures, include=TRUE}

# scatterplot: PAR x month
ggplot(data = parmod, aes(x = date, y = normalized_par)) + 
  geom_point(aes(color = site)) +
  labs(title = "Normalized PAR over time", x = "MAFA canopy cover (%)", y = "Date") +
  theme_bw() 

# boxplot: PAR by month <---- KIT - can you make a figure that shows a boxplot of normPAR for each month? <-- need to change YEAR, MONTH, SITE, to character. 
#ggplot(data = parmod, aes(x = month, y = normalized_par)) +
#  geom_boxplot() +
#  theme_bw() 

# boxplot: PAR by site x treatment
ggplot(data = parmod, aes(x = treatment, y = normalized_par)) +
  geom_boxplot(aes(color = site)) +
  labs(title = "Normalized PAR over time", x = "MAFA canopy cover (%)", y = "Date") +
  theme_bw() 


```


***

## TDR linear mixed-effects model
1. lmer
2. testing the residuals for normality
```{r tdrmod - normality, include = TRUE}

lme.tdrmod <- lmer(soil_moisture ~ treatment + mafacover + site + (1|date) +(1|plot_number),
                   data = tdrmod)

MOD <- lme.tdrmod

# Calculate model's residual and fitted values
F1 <- fitted(MOD)
E1 <- residuals(MOD, "pearson")

# See residuals as a histogram, look for normal distribution
hist(E1)

# Graph residuals against a normal distribution, look for all points to be on 1:1 line
#need to call next two lines together
qqnorm(E1, xlab = "Theoretical Quantiles", ylab = "Pearson Residuals")
qqline(E1)

# Graph residuals vs fitted values to inspect homogeneity of variance, look for a random scattering of plots
# run together 2 lines
plot(x=F1, y=E1, xlab="Fitted Model Values", ylab="Pearson Residuals", main = "Variance")
abline(h = 0) # looking for random scatter

# Check for crazy outliers, look for lines MUCH taller than the rest of data
plot(cooks.distance(MOD))

# Plot a box plot for each factor, look for equal variance amount levels along the zero line
# Box plots need to be assessed for every fixed effect (categorical factor) for your model. 
# Random effects need the qqline just like normality. 
# So, only site and treatment get box plots. 
# MAFA cover is a covariate (continuous variable).
boxplot(E1 ~ treatment, data = tdrmod)
abline(h = 0)

boxplot(E1 ~ site, data = tdrmod)
abline(h = 0)

# Plot each random effect in the model, look for dots along 1:1 line
qqnorm(ranef(MOD)$date[,1], main = "Plot Residuals")
qqline(ranef(MOD)$date[,1])

qqnorm(ranef(MOD)$plot_number[,1], main = "Plot Residuals")
qqline(ranef(MOD)$plot_number[,1])
```
3. statistics
```{r tdrmod - stats, include = TRUE}
# print 
print(lme.tdrmod)
summary(lme.tdrmod)
anova(lme.tdrmod) # A significant p-value indicates that 
# one or more significant differences exist between group means.

# TDR by Site and Treatment
tdrmod_site_treat <- aov(soil_moisture ~ site*treatment, data = tdrmod) 
anova(tdrmod_site_treat)
TukeyHSD(tdrmod_site_treat)
#plot(TukeyHSD(tdrmod_site_treat))

# TDR by Site
tdrmod_site <- aov(soil_moisture ~ site, data = tdrmod) 
anova(tdrmod_site)
TukeyHSD(tdrmod_site)
plot(TukeyHSD(tdrmod_site))


```
4. figures
```{r tdrmod - figures, include = TRUE}

# boxplot: TDR by treatment by site
ggplot(data = tdrmod, aes(x = treatment, y = soil_moisture)) +
  geom_boxplot(aes(fill = site)) +
  theme_bw() 

# boxplot: TDR by site by treatment
ggplot(data = tdrmod, aes(x = site, y = soil_moisture)) +
  geom_boxplot(aes(fill = treatment)) +
  scale_fill_manual(values = c('#F6EECF', '#B09175', '#291611')) +
  labs(title = "Soil Moisture by Site and Treatment", x = "Site", y = "PAR") +
  theme_bw() 

# scatterplot: tdr x mafacover
ggplot(data = tdrmod, aes(x = mafacover, y = soil_moisture)) + 
  geom_point() +
  geom_smooth(method="lm")+
  theme_bw()
```

***

## Non-native cover linear mixed-effects model
1. lmer
2. testing the residuals for normality
```{r nnmod - normality, include = TRUE}

#need to convert df to a tibble
#as_tibble(nnmod)
#print(nn)
#names(nnmod)
#print(nnmod)
#str(nnmod)
lme.nnmod <- lmer(nncover ~ treatment + mafacover + site + (1|date) + (1|plot_number),
                   data = nnmod)

MOD <- lme.nnmod

# Calculate model's residual and fitted values
F1 <- fitted(MOD)
E1 <- residuals(MOD, "pearson")

# See residuals as a histogram, look for normal distribution
hist(E1)

# Graph residuals against a normal distribution, look for all points to be on 1:1 line
#need to call next two lines together
qqnorm(E1, xlab = "Theoretical Quantiles", ylab = "Pearson Residuals")
qqline(E1)

# Graph residuals vs fitted values to inspect homogeneity of variance, look for a random scattering of plots
# run together 2 lines
plot(x=F1, y=E1, xlab="Fitted Model Values", ylab="Pearson Residuals", main = "Variance")
abline(h = 0) # looking for random scatter

# Check for crazy outliers, look for lines MUCH taller than the rest of data
plot(cooks.distance(MOD))

# Plot a box plot for each factor, look for equal variance amount levels along the zero line
# Box plots need to be assessed for every fixed effect (categorical factor) for your model. 
# Random effects need the qqline just like normality. 
# So, only site and treatment get box plots. 
# MAFA cover is a covariate (continuous variable).
boxplot(E1 ~ treatment, data = nnmod)
abline(h = 0)

boxplot(E1 ~ site, data = nnmod)
abline(h = 0)

# Plot each random effect in the model, look for dots along 1:1 line
qqnorm(ranef(MOD)$date[,1], main = "Plot Residuals")
qqline(ranef(MOD)$date[,1])

qqnorm(ranef(MOD)$plot_number[,1], main = "Plot Residuals")
qqline(ranef(MOD)$plot_number[,1])
```
3. statistics
```{r nnmod - stats, include=FALSE}
# print 
print(lme.nnmod)
summary(lme.nnmod)
anova(lme.nnmod) # A significant p-value indicates that one or more significant differences exist between group means.

# non-native cover by treatment
nnmod_treat <- aov(nncover ~ treatment, data = nnmod) 
anova(nnmod_treat)
TukeyHSD(nnmod_treat)
plot(TukeyHSD(nnmod_treat))

# non-native cover by treatment and site
nnmod_treat_site <- aov(nncover ~ treatment*site, data = nnmod) 
anova(nnmod_treat_site)
TukeyHSD(nnmod_treat_site)
plot(TukeyHSD(nnmod_treat_site))
```

4. figures
```{r nnmod - figures, include=TRUE}

# boxplot: nncover x treatment x site
ggplot(data = nnmod, aes(x = treatment, y = nncover)) + 
  geom_boxplot(aes(fill = site)) +
  geom_smooth(method="lm") +
  labs(title = "Non-native cover by Treatment and Site", x = "Treatment", y = "Non-native cover (%)") +
  scale_fill_manual(values = c( "#DCC27A", "#F19B34")) +
  # scale_fill_manual(values = c('#D3E3CA', '#92A587')) +
  theme_bw()

# boxplot: mafacover x site x treatment
ggplot(data = nnmod, aes(x = site, y = nncover)) + 
  geom_boxplot(aes(fill = treatment)) +
 # scale_fill_manual(values = c('#F19B34', '#9F7E75', '#92A587')) +
  scale_fill_manual(values = c('#F6EECF', '#B09175', '#291611')) +
  labs(title = "Non-native cover by Site and Treatment", x = "Site", y = "Non-native cover (%)") +
  geom_smooth(method="lm") +
  theme_bw()

# scatterplot: nncover x mafacover
ggplot(data = nnmod, aes(x = mafacover, y = nncover)) + 
  geom_point() +
  geom_smooth(method="lm")+
  theme_bw()
```

***

# Question 1: Take aways

### PAR

### TDR 
* Soil moisture is greater at Site 1 than at Site 2


### Non-native cover
1. For TDR and nn cover, the significant plots may explain results we see later, but the fact that there are 2-3 plots that are significantly different from each other doesn't tell us the sites or treatments are significantly different from each other. 



# Question 2: How does seedling survival change by PAR, TDR, and mafacover?
Is canopy treatment or canopy cover a stronger predictor of the environmental conditions below the shrub canopy? 

***

# Glossary

_df exploration_
* __names(df)__ will give you the column names in the df

_statistics_
* lmer: Fit Linear Mixed-Effects Models
  + Fit a linear mixed-effects model (LMM) to data, via REML or maximum likelihood.
* letters
  + https://statdoe.com/cld-customisation/
* ANOVA
  + result <- aov(response_variable ~ explanitory_variable, data = df) 
    TukeyHSD(parmod_treatment)
    plot(TukeyHSD(parmod_treatment))
* t-test: compare the means of two groups 
  + with data from two dfs --> t.test(par1$response_variable, par2(dollarsign)response_variable)
  + with data from one df --> t.test(normalized_par ~ year, data = parmod)
  
  
_end of code_

