---
title: "data_prep_2"
author: "Kit Swift"
date: "2022-11-23"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here) # similar to set working directory
library(tidyverse) # data wrangling
library(lubridate) # dates and times
library(plotly)
library(survival)
library(lme4)
```

# Data Prep

## Data structure
* date: character or date
* year: number
* month: number
* plot number: character
* treatment: ordinal ("none", "half", "full") or character or factor (full, half, none)
* site: factor (1,2)
* survival: factor (0,1)
* percent cover: double
* nn_grassheight_cm: double
* 2018_density: integer
* PAR: double
* TDR: double

## MAFA percent cover
```{r mafa_cover prep, include=TRUE}

# load mafa_alive percent cover data as a .csv
cover_import_prep <- read_csv("data/mafa_cover.csv", 
    col_types = cols(
      date = col_character(), 
      year = col_integer(),
      plot_number = col_integer(), 
      MAFA_removal = col_character(),
      cover_type = col_character(),
      cover_percent = col_number(), # question for SHANE - Percent cover as number or integer?
      notes = col_character()
      )
    )


# 2018
mafa_cover_2018 <- cover_import_prep %>% 
  dplyr::select('date', 'plot_number', 'MAFA_removal', 'cover_type', 'cover_percent') %>% 
  mutate(date = mdy(date),
         month = month(date),
         year = year(date)) %>% 
  filter(year %in% c("2018")) %>%
  dplyr::select('plot_number', 'cover_percent', 'year', 'month') %>%  
  mutate(transplanted = 2018) %>% 
  mutate(site = case_when(plot_number <= 18 ~ '1', plot_number >= 19 ~ '2')) 
print(mafa_cover_2018)
# 36 rows
# names(mafa_cover_2018)


# 2019
mafa_cover_2019 <- read.csv(here("data", "mafa_cover.csv")) %>%
    dplyr::select('date', 'plot_number', 'MAFA_removal', 'cover_type', 'cover_percent') %>% 
  mutate(date = mdy(date),
         month = month(date),
         year = year(date)) %>% 
  filter(year %in% c("2019")) %>%
 dplyr::select('plot_number', 'cover_percent', 'year', 'month') %>%  
  mutate(transplanted = 2019) %>% 
  mutate(site = case_when(plot_number <= 18 ~ '1', plot_number >= 19 ~ '2'))
print(mafa_cover_2019)
# 36 rows

# all 
mafa_cover_all <- rbind(mafa_cover_2018, mafa_cover_2019) %>%
  dplyr::select("plot_number", "year", "cover_percent", "site")
print(mafa_cover_all)
write.csv(mafa_cover_all, "data/processed/mafa_cover_all.csv")

# 36 x 3 = 108 rows
```

## Total exotic percent cover
```{r totalexotic_cover prep, include=TRUE}

# load total_exotic percent cover data as a .csv
exoticcover_import <- read_csv("data/totexo_cover.csv", 
    col_types = cols(
      date = col_character(), 
      year = col_character(),
      plot_number = col_integer(), 
      cover_percent = col_number(), # question for SHANE - Percent cover as number or integer?
      notes = col_character()
      )
    )

#  2018
exotic_cover_2018 <- exoticcover_import %>% 
  dplyr::select('date', 'plot_number', 'MAFA_removal', 'cover_type', 'cover_percent') %>% 
  mutate(date = mdy(date),
         month = month(date),
         year = year(date)) %>% 
  filter(year %in% c("2018")) %>%
  dplyr::select('date', 'plot_number', 'MAFA_removal', 'cover_percent', 'year', 'month') %>%  
  mutate(transplanted = 2018) %>% 
  mutate(site = case_when(plot_number <= 18 ~ '1', plot_number >= 19 ~ '2'))
print(exotic_cover_2018)
# 36 rows

# 2019
exotic_cover_2019 <- exoticcover_import %>% 
  dplyr::select('date', 'plot_number', 'MAFA_removal', 'cover_type', 'cover_percent') %>% 
  mutate(date = mdy(date),
         month = month(date),
         year = year(date)) %>% 
  filter(year %in% c("2019")) %>%
  dplyr::select('date', 'plot_number', 'MAFA_removal', 'cover_percent', 'year', 'month') %>%  
  mutate(transplanted = 2019) %>% 
  mutate(site = case_when(plot_number <= 18 ~ '1', plot_number >= 19 ~ '2'))
print(exotic_cover_2019)
# 36 rows

# all 
exotic_cover_all <- rbind(exotic_cover_2018, exotic_cover_2019 )%>%
  dplyr::select("date", "year", "site", "plot_number", "MAFA_removal", "cover_percent")
print(exotic_cover_all)
write.csv(exotic_cover_all, "data/processed/exotic_cover_all.csv")
# 36 x 2 = 72 rows
```

## Non-native grass density and height
```{r nn grass density and height, include=TRUE}

# load density csv
nngrass2018_density_raw <- read_csv("data/2018_nngrass_density.csv", 
    col_types = cols(
      year = col_character(),
      plot_number = col_integer(),
      quad_rep = col_character(), 
      species_code = col_character(),
      status = col_character(), 
      lifeform = col_character(), 
      status_lifeform = col_character(), 
      density_individual_200cm2 = col_integer()
      )
    )

nngrass2018_density <- nngrass2018_density_raw %>% 
    na.omit(nngrass2018_density_raw) %>% 
  group_by(year, plot_number, quad_rep, status_lifeform) %>% 
  summarize(total_density_200cm2 = sum(density_individual_200cm2))
view(nngrass2018_density)

exotic_cover_2018_prepfordensity <- exotic_cover_2018 %>% 
  dplyr::select(plot_number, MAFA_removal, cover_percent, year, site) %>% 
  mutate(year = as.character(year)) %>% 
  rename(exo_coverpercent = cover_percent)

mafa_cover_2018_prepfordensity <- mafa_cover_2018 %>% 
  dplyr::select(plot_number, cover_percent, year, site) %>% 
  mutate(year = as.character(year)) %>% 
  rename(MAFA_coverpercent = cover_percent)
  

nngrass2018_cover_density <- nngrass2018_density %>% 
  left_join(exotic_cover_2018_prepfordensity, by = c("year", "plot_number")) %>% 
  left_join(mafa_cover_2018_prepfordensity, by = c("year", "plot_number", "site")) %>% 
    mutate(plot_number = as.character(plot_number)) 

view(nngrass2018_cover_density)
write.csv(nngrass2018_cover_density, "data/processed/nngrass2018_cover_density.csv")


# load nn grass height csv
nngrass2018_height_raw <- read_csv("data/2018_nngrass_height.csv", 
    col_types = cols(
      date = col_character(),
      year = col_character(),
      plot_number = col_integer(),
      treatment = col_character(),
      quad_rep = col_character(), 
      nn_grassheight_cm = col_number()
      )
    )

nngrass2018_height <- nngrass2018_height_raw %>% 
  dplyr::select("year", "plot_number", "quad_rep", "nn_grassheight_cm") %>% 
  na.omit(nngrass2018_height_raw)
glimpse(nngrass2018_height)

exotic_cover_2018_prepforheight <- exotic_cover_2018 %>% 
  dplyr::select(plot_number, MAFA_removal, cover_percent, year, site) %>% 
  mutate(year = as.character(year)) %>% 
  rename(exo_coverpercent = cover_percent)

mafa_cover_2018_prepforheight <- mafa_cover_2018 %>% 
  dplyr::select(plot_number, cover_percent, year, site) %>% 
  mutate(year = as.character(year)) %>% 
  rename(MAFA_coverpercent = cover_percent)
  
nngrass2018_cover_height <- nngrass2018_height %>% 
  left_join(exotic_cover_2018_prepforheight, by = c("year", "plot_number")) %>% 
  left_join(mafa_cover_2018_prepforheight, by = c("year", "plot_number", "site")) %>% 
    mutate(plot_number = as.character(plot_number)) 

head(nngrass2018_cover_height)
write.csv(nngrass2018_cover_height, "data/processed/nngrass2018_cover_height.csv")

```



## Seedling survival
```{r seedling survivial prep, include=TRUE} 
survival_prep_load <- read.csv("data/survival_data.csv")
str(survival_prep_load)
#, 
#    col_types = cols(
#      date = col_character(), 
#      plot_number = col_character(), 
#      MAFA_removal = col_character(),
#     location = col_character(),
#      species = col_character(),
#      #individual = col_character(),
#      stem_height_cm = col_number(),
#      width_1_cm = col_number(),
#      width_2_cm = col_number(),
#      #condition = col_character,
#      alive = col_character(),
#      dead = col_character(),
#      notes = col_character()
#      )
#    )

survival_prep <- survival_prep_load %>% 
  dplyr::select(date, plot_number, MAFA_removal, species, stem_height_cm, width_1_cm, width_2_cm, alive) %>%
 # mutate_all(na_if,alive,"") %>% # fills blank cells with NA
    rename(treatment = MAFA_removal, 
           height = stem_height_cm, 
           width1 = width_1_cm, 
           width2 = width_2_cm
           ) %>% 
  mutate(date = mdy(date),
         month = month(date),
         year = year(date),
         treatment = str_remove(treatment, "removal")
         ) %>% 
  group_by(year, month, plot_number, treatment, species, date, height, width1, width2) %>%
  summarise(alive = sum(alive, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(
    month = as.integer(month),
    year = as.integer(year), 
    height = as.numeric(height), 
    width1 = as.numeric(width1),
    width2 = as.numeric(width2)
  ) %>%  
  mutate(site = case_when(plot_number <= 18 ~ '1', plot_number >= 19 ~ '2'))
      
print(survival_prep)
         
#   mutate(treatment = case_when(.$treatment == "no" ~ "none", TRUE ~ survival_prep_load$treatment)) 
# treatment = case_when(plot_number == "5" ~ "half", TRUE ~ survival_prep_1$treatment))
 
```
```{r survival months since transplanting, include = TRUE}
#Setting the month to be a count up from jan 2018
month_0 <- c("2018-01-30", "2018-02-09", "2018-02-10")
month_1 <- c("2018-02-24", "2018-02-23", "2018-03-03")

#if else statement to create months that count up from month transplanted
for (i in 1:length(survival_prep$date)) {
  if(grepl(month_0[1], survival_prep$date[i])||grepl(month_0[2], survival_prep$date[i])||grepl(month_0[3], survival_prep$date[i])){
    survival_prep$month_count[i] <- 0; survival_prep$transplanted[i] <- 2018} 
  else{
    if(grepl(month_1[1], survival_prep$date[i])||grepl(month_1[2], survival_prep$date[i])||grepl(month_1[3], survival_prep$date[i])){
      survival_prep$month_count[i] <- 1; survival_prep$transplanted[i] <- 2018} 
    else{
      if(survival_prep$year[i] == 2018){
      survival_prep$month_count[i] <- survival_prep$month[i] - 1; survival_prep$transplanted[i] <- 2018}
      else{
        if(survival_prep$year[i] == 2019){
        survival_prep$month_count[i] <- survival_prep$month[i] + 11; survival_prep$transplanted[i] <- 2019}
        else{
          if(survival_prep$year[i] == 2020){
          survival_prep$month_count[i] <- survival_prep$month[i] + 23; survival_prep$transplanted[i] <- 2019}
          else{
            if(survival_prep$year[i] == 2021){
            survival_prep$month_count[i] <- survival_prep$month[i] + 35; survival_prep$transplanted[i] <- 2019}
            else{
              survival_prep$month_count[i] <- FALSE
            }}}}}}}
print(survival_prep)

survival <- subset(survival_prep, select = -date)

print(survival) # will only print the first 100 pages
write.csv(survival, "data/processed/survival.csv")
# ignore warning message
```

```{r test month count, inclued = FALSE}
#month_count_0 <- filter(survival, month_count == 0)
#print(month_count_0)

#month_count_1 <- filter(survival, month_count == 1)
#print(month_count_1)

#month_count_12 <- filter(survival, month_count == 12)
#print(month_count_12)
```

## Seedling mortality
```{r seedling mortality prep, include=TRUE}
mortality_prep_load <- read.csv(here("data", "survival_data.csv")) 

mortality_prep <- mortality_prep_load %>% 
  dplyr::select(date, plot_number, MAFA_removal, species, dead, stem_height_cm, width_1_cm, width_2_cm) %>% 
  rename(treatment = MAFA_removal, 
         height = stem_height_cm,
         width1 = width_1_cm,
         width2 = width_2_cm) %>% 
  mutate(date = mdy(date),
         month = month(date),
         year = year(date),
         treatment = str_remove(treatment, "removal")
         ) %>% 
  group_by(year, month, plot_number, treatment, species, height, width1, width2, date) %>%
  summarise(dead = sum(dead, na.rm = TRUE)) %>% 
    ungroup() %>% 
   mutate(
    month = as.integer(month),
    year = as.integer(year), 
    height = as.numeric(height),
    width1 = as.numeric(width1),
    width2 = as.numeric(width2)
  ) %>%  
  mutate(site = case_when(plot_number <= 18 ~ '1', plot_number >= 19 ~ '2'))
print(mortality_prep)
```

```{r mortality months since transplanting, include = TRUE}

#if else statement to create months that count up from month transplanted
for (i in 1:length(mortality_prep$date)) {
  if(grepl(month_0[1], mortality_prep$date[i])||grepl(month_0[2], mortality_prep$date[i])||grepl(month_0[3], mortality_prep$date[i])){
    mortality_prep$month_count[i] <- 0; mortality_prep$transplanted[i] <- 2018} 
  else{
    if(grepl(month_1[1], mortality_prep$date[i])||grepl(month_1[2], mortality_prep$date[i])||grepl(month_1[3], mortality_prep$date[i])){
      mortality_prep$month_count[i] <- 1; mortality_prep$transplanted[i] <- 2018} 
    else{
      if(mortality_prep$year[i] == 2018){
      mortality_prep$month_count[i] <- mortality_prep$month[i] - 1; mortality_prep$transplanted[i] <- 2018}
      else{
        if(mortality_prep$year[i] == 2019){
        mortality_prep$month_count[i] <- mortality_prep$month[i] + 11; mortality_prep$transplanted[i] <- 2019}
        else{
          if(mortality_prep$year[i] == 2020){
          mortality_prep$month_count[i] <- mortality_prep$month[i] + 23; mortality_prep$transplanted[i] <- 2019}
          else{
            if(mortality_prep$year[i] == 2021){
            mortality_prep$month_count[i] <- mortality_prep$month[i] + 35; mortality_prep$transplanted[i] <- 2019}
            else{
              mortality_prep$month_count[i] <- FALSE
            }}}}}}}

print(mortality_prep)

mortality <- subset(mortality_prep, select = -date)

write.csv(mortality, "data/processed/mortality.csv")
# mortality is the inverse of survival

```

## Seedling survival + mortality in one simple package
```{r seedling survivial and mortality prep, include=TRUE} 
survmort_prep <- read.csv(here("data", "survival_data.csv")) %>%
  dplyr::select(date, plot_number, MAFA_removal, species, alive, dead, stem_height_cm, width_1_cm, width_2_cm) %>%
  rename(treatment = MAFA_removal,
         height = 
          ) %>%
  mutate(date = mdy(date),
         month = month(date),
         year = year(date),
         treatment = str_remove(treatment, "removal")
        ) %>%
  group_by(year, month, plot_number, treatment, species, date) %>%
  summarise(alive = sum(alive, na.rm = TRUE), dead = sum(dead, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(
    month = as.integer(month),
    year = as.integer(year)
  ) %>%
  mutate(site = case_when(plot_number <= 18 ~ '1', plot_number >= 19 ~ '2'))
print(survmort_prep)
```

```{r survmort months since transplanting, include = TRUE}
#Setting the month to be a count up from jan 2018
month_0 <- c("2018-01-30", "2018-02-09", "2018-02-10")
month_1 <- c("2018-02-24", "2018-02-23", "2018-03-03")

#if else statement to create months that count up from month transplanted
for (i in 1:length(survmort_prep$date)) {
  if(grepl(month_0[1], survmort_prep$date[i])||grepl(month_0[2], survmort_prep$date[i])||grepl(month_0[3], survmort_prep$date[i])){
    survmort_prep$month_count[i] <- 0; survmort_prep$transplanted[i] <- 2018; survmort_prep$month[i] <- 1}
  else{
    if(grepl(month_1[1], survmort_prep$date[i])||grepl(month_1[2], survmort_prep$date[i])||grepl(month_1[3], survmort_prep$date[i])){
      survmort_prep$month_count[i] <- 1; survmort_prep$transplanted[i] <- 2018; survmort_prep$month[i] <- 2}
    else{
      if(survmort_prep$year[i] == 2018){
      survmort_prep$month_count[i] <- survmort_prep$month[i] - 1; survmort_prep$transplanted[i] <- 2018}
      else{
        if(survmort_prep$year[i] == 2019){
        survmort_prep$month_count[i] <- survmort_prep$month[i] + 11; survmort_prep$transplanted[i] <- 2019}
        else{
          if(survmort_prep$year[i] == 2020){
          survmort_prep$month_count[i] <- survmort_prep$month[i] + 23; survmort_prep$transplanted[i] <- 2019}
          else{
            if(survmort_prep$year[i] == 2021){
            survmort_prep$month_count[i] <- survmort_prep$month[i] + 35; survmort_prep$transplanted[i] <- 2019}
            else{
              survmort_prep$month_count[i] <- FALSE
            }}}}}}}
print(survmort_prep)

survmort <- subset(survmort_prep, select = -date)

print(survmort) # will only print the first 100 pages
write.csv(survmort, "data/processed/survmort.csv")
# ignore warning message
```
## PAR - sunshine
```{r par, include=TRUE}
par_load <- read.csv(here("data", "par_data.csv")) 

par_load$date <- as.character(par_load$date)

par <- par_load %>% 
  dplyr::select(date, plot, weeding, readingno, ambient_PAR_1, below_PAR_1) %>% 
  rename(plot_number = plot) %>% 
  rename(treatment = weeding) %>% 
  mutate(date = case_when(date == "11-Sep-18" ~ "9/11/2018",
                          date == "30-Jan-19"~"1/30/2019",
                          date == "25-Jan-19"~"1/25/2019",
                          date == "24-Mar-19"~"3/24/2019",
                          date == "24-May-19" ~ "5/24/2019",
                          date == "19-Jul-19"~ "7/19/2019",
                          date == "9-Sep-19" ~ "9/09/2019",
                          date == "16-Nov-19" ~ "11/16/2019",
                          TRUE ~ date),
         plot_number = str_remove(plot_number, "ENH0"),
         plot_number = str_remove(plot_number, "ENH"),
         plot_number = as.integer(plot_number),
         treatment = str_remove(treatment, "removal"),
         date = mdy(date),
         site = case_when(plot_number <= 18 ~ '1', plot_number >= 19 ~ '2')) %>% 
  group_by(date, plot_number, site, treatment) %>% 
  summarize(ambient_par = mean(ambient_PAR_1),
            below_par = mean(below_PAR_1)) %>% 
  ungroup() %>% 
  mutate(normalized_par = below_par/ambient_par,
         month = month(date),
         year = year(date),
         year = as.integer(year),
         month = as.integer(month)) %>% 
  dplyr::select(date, year, month, site, plot_number, treatment, normalized_par)

print(par)
write.csv(par, "data/processed/par.csv")

```

## TDR - soil moisture
```{r tdr, include=True}
tdr <- read.csv(here("data", "tdr_data.csv")) %>%
  dplyr::select(date, plot, weeding, reading, TDR.value) %>%
  rename(plot_number = plot,
         soil_moisture = TDR.value,
         treatment = weeding) %>%
  mutate(plot_number = str_remove(plot_number, "ENH0"),
         plot_number = str_remove(plot_number, "ENH"),
         plot_number = as.numeric(plot_number),
         treatment = str_remove(treatment, "removal"),
         date = dmy(date),
         site = case_when(plot_number <= 18 ~ '1', plot_number >= 19 ~ '2'),
         treatment = str_replace(treatment, "no", "none")) %>%
  group_by(date, plot_number, site, treatment) %>%
  summarize(soil_moisture = mean(soil_moisture, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(month = month(date),
         year = year(date)) %>%
  #select(!date) %>%
  dplyr::select(date, year, month, site, plot_number, treatment, soil_moisture) %>%
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    plot_number = as.integer(plot_number)
    )


print(tdr)
write.csv(tdr, "data/processed/tdr.csv")
```


# Single Doc Join
```{r main_join, include=FALSE}

surv_month <- rep(1:12, each = 36, 4)
surv_year <- rep(2018:2021, each = 432)
surv_plot <- rep(seq(1,36, by = 1), 48)
surv_monthc <- rep(0:47, each = 36)
surv_hold <- data.frame(surv_plot, surv_year, surv_month, surv_monthc) %>%
   rename(month = surv_month,
          year = surv_year,
          plot_number = surv_plot,
          month_count = surv_monthc) %>% 
  mutate(year = as.integer(year))

surv_all <- surv_hold %>%
  left_join(survmort, by = c("month_count", "plot_number"), na.rm = F)  %>%
  mutate(month = coalesce(month.y, month.x),
         year = coalesce(year.x, year.y)) %>%
  left_join(par, by = c("month", "year", "plot_number"), na.rm = F)  %>%
  left_join(tdr, by = c("month", "year", "plot_number"), na.rm = F) %>%
  mutate(site = coalesce(site, site.y, site.x),
         treatment = coalesce(treatment.x, treatment.y, treatment)) %>%
  drop_na(site) %>%
  left_join(mafa_cover_all, by = c("plot_number", "site", "year")) %>%
  left_join(exotic_cover_all, by = c("plot_number", "site", "year"), suffix = c(".native", ".nonnative")) %>%
  dplyr::select("year","month", "month_count", "site", "plot_number", "treatment", "species", "alive", "dead", "normalized_par", "soil_moisture", "cover_percent.native", "cover_percent.nonnative") %>%
  rename(exotic_cover = cover_percent.nonnative,
         native_cover = cover_percent.native)

write.csv(surv_all, "data//processed/surv_all.csv")
surv_months <- unique(surv_all$month_count)
surv_months
# [1]  0  1  3  4  8 11 12 14 15 16 17 18 20 22 26 27 28 33 37 39
```





