---
title: "Piru_enhancement_seedlingsurvival"
author: "Stephanie Ma Lucero"
date: "`r Sys.Date()`"
output: html_document
---

# Load libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(here) # similar to set working directory
library(tidyverse) # data wrangling
```

# Load colors for graphs

```{r colors, include=TRUE}
devtools::install_github("an-bui/calecopal")

library(calecopal)
# all palettes
names(cal_palettes)

# Site colors: 1, 2 
cal_palette(name = "chaparral1", n = 5, type = "continuous") # [1] "#DCC27A" "#F19B34"
site1color <- "#DCC27A"
site2color <- "#F19B34"
sitecolors <- c(site1color, site2color)

# Treatment colors: Full, half, none 
cal_palette(name = "chaparral3", n = 5, type = "continuous") # [1] "#D3E3CA", "#92A587", "#2F3525"
cal_palette(name = "desert", n = 5, type = "continuous") # [1] "#F6EECF", "#B09175", "#291611"
treatment1color <- "#F6EECF"
treatment2color <- "#B09175"
treatment3color <- "#291611"
treatmentcolors <- c(treatment1color, treatment2color, treatment3color)

```

# Question 2: How does seedling survival change with PAR, TDR, % MAFA cover, and species?

Notes:
* __response variable__: The variable you want to test. (e.g.,. PAR, TDR, nncover)
* __explanatory variables__: The variables that can explain the outcome. (e.g., treatment, mafacover, site)
* __data frame__: Include all the data you'll be working with. (e.g., combine par and mafa_cover_all)
* __formula__: a formula is a two-sided linear formula object describing both the _fixed-effects_ and _random effects_ part of the model, with the response on the left of a ~ operator and the terms, separated by + operators, on the right. 
* __random effects__: Random-effects terms are distinguished by vertical bars __(|)__ separating expressions for design matrices from grouping factors. Two vertical bars __(||)__ can be used to specify multiple uncorrelated random effects for the same grouping variable. 
  + Because of the way it is implemented, the ||-syntax works only for design matrices containing numeric (continuous) predictors; to fit models with independent categorical effects, see dummy or the lmer_alt function from the afex package.)
  
  
  
  
  
  
  