---
title: "TDR"
author: "Stephanie Ma Lucero"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here) # similar to set working directory
library(tidyverse) # data wrangling
library(lubridate) # dates and times
library(ggplot2)
```

```{r view tdr, include=TRUE}
print(tdr)
```
Testing the TDR dataset for normality.
```{r tdr test for normality}
tdr_normality <- ggplot(tdr) + geom_histogram(aes(x = soil_moisture))
tdr_normality

# it's not normally distributed
```

Scatter Plot: TDR x % MAFA cover
# Select desired data from MONTH and YEAR 
#Join MAFA percent cover and TDR (soilmoisture)
```{r, include=FALSE}
# TDR 
# 2018 april, dec
# 2019 march
# 2020 march, april, may
# 
# MAFA cover:
# 2017	8-9
# 2019	5-6			
```

# 2019 
##Join 2019 march TDR data with 2019 MAFA cover data by plot number and month
##Purpose: MAFA cover as an explanatory variable of TDR (soil moisture)
```{r 2019 tdr_percentcover join, include=TRUE}
mafa_cover.2019 <- mafa_cover %>% 
  filter(year == "2019")

tdr_mafacover.2019 <- tdr %>% 
  filter(year == "2019") %>% 
  filter(month == "3") %>% 
  # TDR: 2019 march
  left_join(mafa_cover.2019, by = c("plot_number")) %>% 
  rename(year.TDR = year.x) %>% 
  rename(month.TDR = month.x) %>% 
  rename(year.MAFAcover = year.y) %>% 
  rename(month.MAFAcover = month.y) %>% 
  select(year.TDR, month.TDR, plot_number, soil_moisture, cover_percent, treatment.y) %>% 
  drop_na() %>% 
  rename(treatment = treatment.y)

print(tdr_mafacover.2019)
```

##TDR scatter plot: March 2019
``` {r tdr scatterplot march 2019 and by mafa cover, include=TRUE}
tdr_cover_scatter_2019 <- ggplot(
  data = tdr_mafacover.2019,
  aes(x = cover_percent, y = soil_moisture, color = treatment)) +
  geom_point(size=2, shape=16) +
 # scale_color_manual(values=c('#D3E3CA', '#92A587', '#2F3525', '#9F7E75', '#B9B994', '#F19B34')) +
  theme(legend.position="top") +
  theme_classic() +
  labs(title="Soil Moisture by Treatment in March 2019",x="Canopy Cover (%)", y = "Soil Moisture (%)") +
  geom_smooth()
print(tdr_cover_scatter_2019)

```

# 2020
##Join 2020 march TDR data with 2019 MAFA cover data by plot number and month
##Purpose: MAFA cover as an explanatory variable of TDR (soil moisture)
```{r 2020 tdr_percentcover join, include=TRUE}

tdr_mafacover.2020 <- tdr %>% 
  filter(year == "2020") %>% 
  filter(month == "3") %>%  # <-- CHANGE MONTH
  # 2020 march, april, may
  left_join(mafa_cover.2019, by = c("plot_number")) %>% 
  rename(year.TDR = year.x) %>% 
  rename(month.TDR = month.x) %>% 
  rename(year.MAFAcover = year.y) %>% 
  rename(month.MAFAcover = month.y) %>% 
  select(year.TDR, month.TDR, plot_number, soil_moisture, cover_percent, treatment.y) %>% 
  drop_na() %>% 
  rename(treatment = treatment.y)

print(tdr_mafacover.2020)
```

##TDR scatter plot: March 2020
``` {r tdr scatterplot march 2020 and by mafa cover, include=TRUE}
tdr_cover_scatter_2020 <- ggplot(
  data = tdr_mafacover.2020,
  aes(x = cover_percent, y = soil_moisture, color = treatment)) +
  geom_point(size=2, shape=16) +
 # scale_color_manual(values=c('#D3E3CA', '#92A587', '#2F3525', '#9F7E75', '#B9B994', '#F19B34')) +
  theme(legend.position="top") +
  theme_classic() +
  labs(title="Soil Moisture by Treatment in March 2020",x="Canopy Cover (%)", y = "Soil Moisture (%)") +
  geom_smooth()

print(tdr_cover_scatter_2020)

```
