---
title: "Licor_data"
author: "Stephanie Ma Lucero"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Install: 
+ packages
```{r install packages}
install.packages("ggplot2")
install.packages("tidyverse")
devtools::install_github("an-bui/calecopal") # California color pallets https://github.com/an-bui/calecopal
```
+ libraries
```{r install libraries}
library(tidyverse) # includes ggplot2, dplyr, tidyr, readr, purr, tibble, stringr, forcats
library(ggplot2) # data visualization
library(readr)
library(dplyr)
library(vegan)

library(calecopal)
names(cal_palettes) # all palettes
```

2. Load csv file.
```{r load csv}
licor <- read.csv("~/Desktop/dataforR.csv", header = TRUE)
```

3. Organize the data
```{r}
dim(licor)
head(licor)
# view(licor) # to view full data table

# convert dataframe to tibble
licor <- licor %>% 
  as_tibble(licor) 

class(licor) # now a tibble
# [1] "tbl_df"     "tbl"        "data.frame"
```
4. Create new column with spp_rep
```{r creat spp_rep column}
licor$spp_rep<- paste(licor$species, licor$rep, sep="_") # combine site and rep to one column
head(licor) # look at new column spp_rep, last column
```
5. Check for normality 
+ all data
```{r select 2000 PAR}
licor_norm <- ggplot(licor, aes(x = Photo)) + 
  geom_histogram(binwidth = 1)
#  geom_histogram(binsize = 15)
print(licor_norm)
# data not normally distributed.
```
+2000 PAR
```{r select 2000 PAR}
licor_2000 <- licor %>% # call tibble
  filter(PAR > 1900 & PAR < 2100) %>%  # select 2000 data points
  group_by(spp_rep)

print(licor_2000) # there should be 12 rows of data
```
```{r test for normality 2000 PAR} 
licor_2000_norm <- ggplot(licor_2000, aes(x = Photo)) + 
  geom_histogram(binwidth = 1)
print(licor_2000_norm)
# 2000 data not normally distributed.
```
+ 1000
```{r select 2000 PAR}
licor_1000 <- licor %>% # call tibble
  filter(PAR > 900 & PAR < 1100) %>%  # select 2000 data points
  group_by(spp_rep)

print(licor_1000) # there should be 12 rows of data
```
```{r test for normality 1000 PAR} 
licor_1000_norm <- ggplot(licor_1000, aes(x = Photo)) + 
  geom_histogram(binwidth = 1)
print(licor_1000_norm)
# 1000 data not normally distributed.
```
If data not normally distributed, let's use the...






If there are no outliers in the dataset, we can use the Kruskal Wallis rank test to determine significance of abrich_totalrich (Total Species Richness aboveground).

Testing for outliers.
```{r abrich total rich testing for outliers}
ggplot(abrich_totalrich) + # no outliers in this dataset
  aes(x = "All data points", y = total_species_richness) +
  geom_boxplot(fill = "#0c4c8a",
               outlier.shape = 9) + #shape 9 = diamond with intersecting lines
  theme_bw()
```

Since abrich_totalrich is not normally distributed and there are no outliers, let's use the rank-based Kruskal Wallis test to determine significance of above ground Total Species Richness.
```{r abrich kruskal.test}
#rank based - Kurskal-Wallis
# Josie Lesage also suggests using the beta ray distribution - good for % cover distribution

abrich_totalrich_kw <- kruskal.test(total_species_richness ~ standtype, data = abrich_totalrich) 
abrich_totalrich_kw
# Kruskal-Wallis chi-squared = 4.1256, df = 2, p-value = 0.1271
```

Now let's calculate the **mean** and **standard errors** of the
    total species richness from each sampling location by STANDTYPE (belt and point).
```{r abrich calculating total species richness mean and SD by stand}

abrich_totalrich_standemeanSD <- abrich_totalrich %>%
  group_by(standtype) %>% # to graph by standtype
  summarise_at(vars(total_species_richness), funs(mean,sd)) 
 
print(abrich_totalrich_standemeanSD)
```
We can also calculate the mean and SE of abrich_totalrich by SITE. 
```{r abrich calculating total species richness mean and SD by site}

abrich_totalrich_sitemeanSD <- abrich_totalrich %>%
  group_by(standtype, site) %>% # to graph all 7 sites
  summarise_at(vars(total_species_richness), funs(mean,sd)) 

print(abrich_totalrich_sitemeanSD)
```